---
title: 'Recitation #2: Agenda'
author: "Tyler Simko"
date: "9/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# normally in a .Rmd file, you'll load packages in this first chunk

library(PPBDS.data)

```

I have played a sneaky trick. By placing this template file on GitHub, we accomplished three goals at once:

1. You practiced cloning a repository from GitHub! 
2. You now know the agenda for today's recitation.
3. You now have a handy place to take notes and practice code for today. 

A fourth, smaller, goal is that you can see some cool formatting tricks in RMarkdown (like the numbered list above). You can find a bunch more by looking at this [list of cheat sheets](https://rstudio.com/resources/cheatsheets/).

## Census Maps

A very common data source for many of you will be the US Census. You can get a ton of demographic information at multiple levels, from the state to the neighborhood level. We're going to walk through [a few sections of the textbook together](https://davidkane9.github.io/PPBDS/maps.html). 

## Spot the bugs!

Error messages are scary but - believe it or not - someone was trying to help you when they wrote it. Most of the time, if you carefully read the error message, you can figure out a rough idea of where your bug is. This is especially true once you start to recognize common patterns in bugs. 

These were very common mistakes from Tutorial 2. Let's find the bugs together:

```{r, eval=FALSE}

#####################################################
# 1. I used position_dodge! 
#    Why are these bars not next to each other?
#####################################################

sps %>%
  filter(education %in% c("preschool", "secondary", 
                          "high school", "college")) %>%
  select(-c(health_exp_1m, t2_health_exp_1m)) %>%
  mutate(change_in_expenses = t2_health_exp_3m - health_exp_3m) %>%
  group_by(treatment, education) %>%
  summarize(mean_change_in_expenses = mean(change_in_expenses)) %>% 
  mutate(as.factor(treatment)) %>%
  ggplot(mapping = aes(x = fct_reorder(education, mean_change_in_expenses), 
                       y = mean_change_in_expenses, fill = treatment)) +
    geom_col(position = position_dodge(preserve = "single"))

#####################################################
# 2. Attempt to replicate an object of type 
#####################################################

shaming %>%
  select(primary_04, primary_06, treatment) %>%
  mutate(primary_04 = if_else(condition = 
                                str_detect(primary_04, "Yes"),
                                true = 1L,
                                false = 0L)) %>%
  pivot_longer(cols = c(primary_04, primary_06), 
               names_to = "year", values_to = "voted") %>%
  group_by(treatment, year, voted) %>%
  summarize(num_voters = n(), .groups = "drop") %>%
  mutate(voted = if_else(condition = voted == 1,
                         true = "did_vote",
                         false = "did_not_vote")) %>%
  pivot_wider(names_from = voted, values_from = num_voters)  %>%
  mutate(pct_voted = did_vote / (did_vote + did_not_vote)) %>% 
  ggplot(mapping = aes(x = fct_reorder(treatment, pct_voted), 
                       y = pct_voted, fill = year)) +
    geom_col(position = position_dodge(preserve = "single")) +
    coord_cartesian(ylim(0.2, 0.45))

#####################################################
# 3. Object not found
#####################################################

# Example 1

cces %>%
  mutate(Age = Age / 10)

# Example 2

sps %>% 
  mutate(treatment_string = if_else(treatment == 1, "treatment", "control"))

ggplot(sps, aes(x = age, y = health_exp_3m, fill = treatment_string)) + 
  geom_col()

# Example 3

sps_edited <- sps %>% 
  mutate(treatment_string = if_else(treatment == 1, "treatment", "control"))

ggplot(sps, aes(x = age, y = health_exp_3m, fill = treatment_string)) + 
  geom_col()

# Example 4

qscores %>% 
  mutate(above_50 = if_else(enrollment > 50, "Large Class", "Small Class")) %>%
  ggplot(data = qscores, mapping = aes(x = hours, y = rating, color = above_50)) + 
    geom_point()

# Example 5

qscores %>% 
    mutate(above_50 = if_else(enrollment > 50, "Large Class", "Small Class")) %>%
    ggplot(qscores, aes(x = hours, y = rating, color = above_50)) + 
    geom_point()

# Example 6: this one is hard!!!

qscores %>% 
    mutate(above_50 = if_else(enrollment > 50, "Large Class", "Small Class")) %>%
    ggplot(qscores, mapping = aes(x = hours, y = rating, color = above_50)) + 
    geom_point()

#####################################################
# 4. Unexpected input in
#####################################################

nes %>%
    select(year, education, pres_appr, ideology, voted) %>%
    filter(education == c(“Some Highschool”, “Highschool”, 
                           “Some College”, “Adv. Degree”))
```

## Some Practice

Here are some topics that students asked to cover in Recitation:

#### 1. joins (full, inner, left, right, etc.)

#### 2. plot themes

#### 3. if_else

#### 4. pivots

#### 5. str_detect